Nail: A practical interface generator for data formats

### 2014-05-02 paper/osdi14/eval.tex
Motivation figure

\cc{unzip}
  & 220
  & 1,600
  & (Info-Zip) \\
\bottomrule
\end{tabular}
\caption{Comparison of code size for three applications written in Nail,
         and several existing alternative implementations.}
\label{fig:effort}
\end{figure}

@@

Our DNS server parses a zone file, listens to incoming DNS requests,
parses them, and generates appropriate responses.  The DNS server is
implemented in 183 lines of C, together with 48 lines of Nail grammar
and 64 lines of C code implementing stream transforms for DNS label
compression.  In comparison, Hammer~\cite{hammer-parser} ships with a toy
DNS server that responds to any valid DNS query with a CNAME record to the
domain ``spargelze.it''.  Their server consists of 683 lines of C, mostly
custom validators, semantic actions, and data structure definitions, with
only 52 lines of code defining the grammar with Hammer's combinators.
Their DNS server does not implement label compression, zone files, etc.
From this, we conclude that Nail leads to much more compact code for
dealing with DNS packet formats.

>>>>

\cc{unzip}
  & 220
  & 1,600
  & (Info-Zip) \\
\bottomrule
\end{tabular}
\caption{Comparison of code size for three applications written in Nail,and several existing alternative implementations.}
\label{fig:effort}
\end{figure}

@@

Our DNS server parses a zone file, listens to incoming DNS requests,
parses them, and generates appropriate responses.  The DNS server is
implemented in 183 lines of C, together with 48 lines of Nail grammar
and 64 lines of C code implementing stream transforms for DNS label
compression.  In comparison, Hammer~\cite{hammer-parser} ships with a toy
DNS server that responds to any valid DNS query with a CNAME record to the
domain ``spargelze.it''.  Their server consists of 683 lines of C, mostly
custom validators, semantic actions, and data structure definitions, with
 52 lines of code defining the grammar with Hammer's combinators.
Their DNS server does not implement label compression, zone files, etc.
From this, we conclude that Nail leads to much more compact code for
dealing with DNS packet formats.

### 2014-05-02 paper/osdi14/motivation.tex
Motivation figure

\paragraph{Vulnerability mitigation.}

Input processing vulnerabilities fall into two broad classes.  The first
class is memory safety bugs, such as buffer overflows, which allow an
adversary to corrupt the application's memory using specially crafted
inputs.  These mistakes arise in lower-level languages that do not
provide memory safety guarantees such as C, and can be partially
mitigated by a wide range of techniques, for example static analysis,
dynamic instrumentation, and address space layout randomization, that
make it more difficult for an adversary to exploit these bugs.
Nail helps developers of lower-level languages avoid these bugs
in the first place.

The second class is logic errors, where application code misinterprets
input data. Safe languages and exploit mitigation technologies do not
help against such vulnerabilities. This can lead to serious security
consequences when two systems disagree on the meaning of a network
packet or a signed message.  For example, the XNU kernel in iOS and the
user-mode code-signing verifier interpret executable metadata differently,
so the code signature sees different bytes at a virtual address than the
executable that runs~\cite{evaders6}.  Android's code signing system also
had a vulnerability due to parsing ambiguities~\cite{saurik-masterkey},
and parsing ambiguities even contributed to attacks on the X.509 protocol
underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  Logic errors can also
cause memory corruption bugs as well, for example when a protocol stores
the size of a buffer in two places and the implementation does not check
their consistency. These mistakes are highly application-specific, and
are difficult to mitigate using existing techniques, and these mistakes
can occur even in high-level languages that guarantee memory safety. By
allowing developers to specify their data format just once, Nail avoids
logic errors and inconsistencies in parsing and output generation.

>>>>

To motivate the need for Nail, this section presents a case study of
vulnerabilities due to ad-hoc input parsing and output generation.

\paragraph{Parsing bugs} Two recent high profile security vulnerabilities in mobile operating systems are due to improper
input recognition, without using any memory corruption. In both cases, when the vulnerabilities were
fixed, a similar flaw was exposed immediately afterwards, showing the need for a different approach
to input handling that eliminates those vulnerabilities by design.

The Evasi0n jailbreak for iOS 6~\cite{evaders6} relies on the XNU kernel and
user-mode code-signing verifier interpreting executable metadata differently, so the code signature
sees different bytes at a virtual address than kernel maps into the process. The next version of iOS
added an explicit check for this particular metadata inconsistency. However, because parsing and
processing of the input data is still mixed, the jailbreakers could set a flag that re-introduced
the ambiguity after the check, but before signatures are verified~\cite{geohot-evasion}, which
allowed iOS 7 to be jailbroken.

Implementations of the X.509 format for SSL certificates were also shown to parse various ASN.1
structures, such as length fields, strings containing invalid characters and repeated fields with
conflicting values different ~\cite{DBLP:conf/fc/KaminskyPS10}. For example, certificates could be
crafted where the X.509 parsers used certificate authorities and web browsers disagree on the
subject of the certificate, allowing an attacker to get certificates for domains they do not control.

Similarly, the infamous Android master key bug~\cite{saurik-masterkey} completely bypassed Android
security exploited parser inconsistencies
between the ZIP handler that checks signatures for privileged applications and the ZIP
implementation that ultimately extract those files. Thus, privileged application bundles could be
modified to include malicious code without breaking their signatures. Google quickly fixed this
particular parser inconsistency, but another vulnerability, based on a different inconsistency
between the parsers, was quickly disclosed~\cite{saurik-masterkey2}.

Android application bundles are based on ZIP files, so we decided to conduct a systematic studies of
vulnerabilities related to ZIP parsing.  
The ZIP format has been associated with many vulnerabilities, and the
PROTOS Genome project~\cite{protoc-c10-archive} found numerous security vulnerabilities in most
implementations of ZIP and other archive formats that are directly
related to input handling. We extend this study by looking at the CVE database.

\input{fig-motivation}
Figure~\ref{fig:motivation} shows  the 83 vulnerabilities the CVE database~\cite{cve-database} lists
 for the search string ``ZIP''. 16 of these vulnerabilities are related to processing .ZIP archives
 themselves; the rest are either totally unrelated to .ZIP archives or involve applications insecurely using
 the contents of untrusted ZIP files.

These input processing vulnerabilities fall into two broad classes.  The first
class, of occurred 11 times, is memory safety bugs, such as buffer overflows, which allow an
adversary to corrupt the application's memory using specially crafted
inputs.  These mistakes arise in lower-level languages that do not
provide memory safety guarantees such as C, and can be partially
mitigated by a wide range of techniques, for example static analysis,
dynamic instrumentation, and address space layout randomization, that
make it more difficult for an adversary to exploit these bugs.
Nail helps developers of lower-level languages avoid these bugs
in the first place.

The second class, which occurred four times in our study is logic errors, where application code
misinterprets input data. Safe languages and exploit mitigation technologies do not help against
such vulnerabilities. This can lead to serious security consequences when two systems disagree on
the meaning of a network packet or a signed message, as shown by the vulnerabilities we described
before. CVE-2013-0211 shows that logic errors can be the underlying cause of memory corruption, when
one part of a parser interprets a size field as a signed integer and another interprets it as an
unsigned integer. CVE-2013-7338 is a logic error that allows an attacker to craft ZIP files that are
incorrectly extracted or result in applications hangs with applications using a Python ZIP library,
because this library does not check that two fields that contain the size of a file contain the same value.

 These mistakes are highly application-specific, and
are difficult to mitigate using existing techniques, and these mistakes
can occur even in high-level languages that guarantee memory safety. By
allowing developers to specify their data format just once, Nail avoids
logic errors and inconsistencies in parsing and output generation.
e through endless loops
or incorrectly extracted ZIP files.

The fourth vulnerability (CVE-\XXX) is the Android master key bug mentioned above, which results
from one parser interpreting a size field as an unsigned integer, another interpreting the same
field as a signed integer.
